### 2021-07-27 paper/localpsf.tex
edits to impulse response moments section

We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). 
%To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions. 
A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}). We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}). For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
%\begin{itemize}
%\item We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions.
%\item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}). 
%\item We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\psi_x$ (Section \ref{eq:mean_and_covariance_estimation}). 
%\item For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
%\end{itemize}
The complete algorithm for constructing $\mathcal{A}_\text{pc}$ is shown in Algorithm \ref{alg:construct_Atilde}. The method for evaluating kernel entries of $\Phi_\text{pc}$ is shown in Section \ref{sec:eval_matrix_entries}. The method for converting $\mathcal{A}_\text{pc}$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}. The method for modifying the resulting H-matrix to make it symmetric positive semi-definite is shown in Section \ref{sec:make_spd}.

@@

Since $\phi_x$ is nonnegative, it is a scaled probability distribution. Let $\widehat{\phi}_x$ denote the normalized version of $\phi_x$, and let $\mu(x)$ and $\Sigma(x)$ denote the mean and covariance of $\widehat{\phi}_x$, respectively (see Definition \ref{defn:alpha_rho_mu_sigma}). We make the approximation that the support of $\phi_x$ is contained within the ellipsoid
\begin{equation}
\label{eq:support_ellipsoid}
E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
\end{equation}
where $\tau$ is a fixed constant. 
%The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance. 
The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\phi}_x$ (see Figure FIG). The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be, while the smaller $\tau$ is, the more impulse responses we can fit within each batch. In practice we find that $\tau=3$ provides a good balance between these competing goals. 
The fraction of the ``mass'' of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.

In Algorithm \ref{alg:varhpi_mean_cov} we show how to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$ for all points $x$ simultaneously, by applying $\mathcal{A}^T$ to a constant function and a small number of linear and quadratic functions. Definitions required for Algorithm \ref{alg:varhpi_mean_cov} are given in Definition \ref{defn:alpha_rho_mu_sigma}, Definition \ref{defn:C_L_Q}, and Definition \ref{defn:pointwise}. The theoretical basis for Algorithm \ref{alg:varhpi_mean_cov} is shown in Theorem \ref{thm:vol_mean_cov}.

@@

\begin{defn}[Constant, linear, and quadratic functions]	
	\label{defn:C_L_Q}
	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$. We define $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
	\begin{equation*}
	C(x) := 1, \qquad
	L^i(x) := x^i, \qquad
	Q^{ij}(x) = x^i x^j.
	\end{equation*}
\end{defn}

\begin{defn}[Pointwise operations]
	\label{defn:pointwise}
	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
\end{defn}

@@

\begin{theorem}[Correctness of Algorithm \ref{alg:varhpi_mean_cov}]
	\label{thm:vol_mean_cov}
	Algorithm \ref{alg:varhpi_mean_cov} correctly computes $\alpha$, $\mu$, and $\Sigma$.
\end{theorem}

The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}.

@@

Now that we have estimates of the supports of the functions $\phi_x$, we use these estimates, combined with a greedy algorithm to choose sets of sample points, $S_b$. These sample points will be used in Section \ref{sec:get_impulse_response} to form Dirac combs, which are then used to compute impulse responses of $\mathcal{A}$. Starting from a finite set of candidate points $P$, we select batches of sample points, $S_b$, using a greedy algorithm. To build a batch $S_b$, first we initialize $S_b$ as an empty set. Then we select a candidate point $p \in P$ that is the farthest away from all points in previous sample point batches. That is, $p$ is a maximizer of the following optimization problem:
\begin{equation*}
\max_{p \in P} \min_{x \in S_1 \cup \dots \cup S_{b-1}} \|p - x\|.
\end{equation*}
Once $p$ is selected, we remove $p$ from $P$. If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$, and if $p$ is sufficiently far from the boundary, in the sense that $E_p \cap \partial \Omega = \{\}$, then we add $p$ to $S_b$. Otherwise we discard $p$. Recall that $E_p$ and $E_q$ are ellipsoids defined in \eqref{eq:support_ellipsoid}. This process repeats until there are no more points in $P$.  This is detailed in Algorithm \ref{alg:point_choice}. We determine whether $E_p \cap E_q = \{\}$ using the fast ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.

>>>>

To briefly summarize, we extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). 
%To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions. 
A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}). We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}). For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
%\begin{itemize}
%\item We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions.
%\item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}). 
%\item We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\psi_x$ (Section \ref{eq:mean_and_covariance_estimation}). 
%\item For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
%\end{itemize}
The complete algorithm for constructing $\mathcal{A}_\text{pc}$ is shown in Algorithm \ref{alg:construct_Atilde}. The method for evaluating kernel entries of $\Phi_\text{pc}$ is shown in Section \ref{sec:eval_matrix_entries}. The method for converting $\mathcal{A}_\text{pc}$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}. The method for modifying the resulting H-matrix to make it symmetric positive semi-definite is shown in Section \ref{sec:make_spd}.

@@

Because of the non-negative kernel property, for each $x \in \Omega$, the impulse response $\phi_x$ is a non-negative function $y \mapsto \phi_x(y)$. Hence, $\phi_x$ is a scaled probability distribution. Let $\alpha$ denote the scaling factor, let $\widehat{\phi}_x := \phi_x / \alpha(x)$ denote the normalized version of $\phi_x$, and let $\mu(x)$ and $\Sigma(x)$ denote the mean and covariance of $\widehat{\phi}_x$, respectively. We make the approximation that the support of $\phi_x$ is contained within the ellipsoid
\begin{equation}
\label{eq:support_ellipsoid}
E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
\end{equation}
where $\tau$ is a fixed constant. 
%The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance. 
The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\phi}_x$ (see Figure FIG). The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm. The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\phi_x$. However, the cost of our algorithm will depend on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$. We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results. The fraction of the ``mass'' of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region. 
% a good balance between these competing goals. 
%The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be, while the smaller $\tau$ is, the more impulse responses we can fit within each batch.

The ellipsoid $E_x$ is not immediately available. To use $E_x$, we must first compute the scalar $\alpha(x)$, the length-$d$ vector $\mu(x)$, and the $d \times d$ matrix $\Sigma(x)$ (recall $d \in \{1,2,3\}$ is the spatial dimension of $\Omega$). The direct approach to compute these quantities is to apply $\Aop$ to a delta distribution centered at point $x$ to get $\phi_x$ (as per \eqref{eq:impulse_response_delta_action}), then post-process $\phi_x$ to compute these quantities. But this direct approach is computationally infeasible because and it requires applying $\Aop$ to one vector per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points. Fortunately, it is possible to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$, \emph{for all points $x$ simultaneously}, by applying $\mathcal{A}^T$ to one constant function, $d$ linear functions, and $d(d+1)/2$ quadratic functions. We present this efficient method of computing $\alpha$, $\mu$, and $\Sigma$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.

%In Algorithm \ref{alg:varhpi_mean_cov} we show how to compute  Definitions required for Algorithm \ref{alg:varhpi_mean_cov} are given in Definition \ref{defn:alpha_rho_mu_sigma}, Definition \ref{defn:C_L_Q}, and Definition \ref{defn:pointwise}. The theoretical basis for Algorithm \ref{alg:varhpi_mean_cov} is shown in Theorem \ref{thm:vol_mean_cov}.

@@

\begin{theorem}[Impulse response moments]
	\label{thm:vol_mean_cov}
	Let $\alpha:\Omega \rightarrow \mathbb{R}$, $\mu:\Omega \rightarrow \mathbb{R}^d$, and $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$, be the following functions:
	\begin{equation*}
	\alpha(x) := \int_{\Omega} \phi_x(y) dy, \qquad
	\mu(x) := \mathbb{E}(\widehat{\phi}_x), \qquad
	\Sigma(x) := \mathbb{V}(\widehat{\phi}_x),
	\end{equation*}
	where $\mathbb{E}(\widehat{\phi}_x)$ denotes the mean of $\widehat{\phi}_x$, and $\mathbb{V}(\widehat{\phi}_x)$ denotes the variance of $\widehat{\phi}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$. Let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions:
	\begin{equation*}
	C(x) := 1, \qquad
	L^i(x) := x^i, \qquad
	Q^{ij}(x) = x^i x^j.
	\end{equation*}
	Let $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$. The following results hold:
	\begin{align*}
	\alpha =& \left(\mathcal{A}^T C\right)^* \\
	\mu^i =& \left(\mathcal{A}^T L^i\right)^* / \alpha \\
	\Sigma^{ij} =& \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j
	\end{align*}
	 for $i=1,\dots, d$, $j=1,\dots,d$.
\end{theorem}

The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.

%
%\begin{defn}[Constant, linear, and quadratic functions]	
%	\label{defn:C_L_Q}
%	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$. We define $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
%	\begin{equation*}
%	C(x) := 1, \qquad
%	L^i(x) := x^i, \qquad
%	Q^{ij}(x) = x^i x^j.
%	\end{equation*}
%\end{defn}
%
%\begin{defn}[Pointwise operations]
%	\label{defn:pointwise}
%	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
%\end{defn}

@@

@@

Now that we have estimates of the supports of the functions $\phi_x$, we use these estimates, combined with a greedy ellipsoid packing algorithm to choose sets of sample points, $S_b$. These sample points will be used in Section \ref{sec:get_impulse_response} to form Dirac combs, which are then used to compute impulse responses of $\mathcal{A}$. We start with a finite set of candidate points $P$. To build a batch of sample points, $S_b$, first we initialize $S_b$ as an empty set. Then we select the point $p \in P$ that is the farthest away from all points in previous sample point batches. That is, $p$ is a maximizer of the following optimization problem:
\begin{equation*}
\max_{p \in P} \min_{x \in S_1 \cup \dots \cup S_{b-1}} \|p - x\|.
\end{equation*}
Once $p$ is selected, we remove $p$ from $P$. If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$, and if $p$ is sufficiently far from the boundary, in the sense that $E_p \cap \partial \Omega = \{\}$, then we add $p$ to $S_b$. Otherwise we discard $p$. Recall that $E_p$ and $E_q$ are ellipsoids associated with points $p$ and $q$, respectively, as defined in \eqref{eq:support_ellipsoid}. This process repeats until there are no more points in $P$.  This is detailed in Algorithm \ref{alg:point_choice}. We determine whether $E_p \cap E_q = \{\}$ using the fast ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.
